\section{Appendix B:  Study Guide}

\subsection{Chapter 2}
\begin{enumerate}

\item{Can you define an angle?  Can you draw an angle?}

\item{Can you define a radian?  A degree?}

\item{Do you know the symbol for radians?  For degrees?}

\item{Can you convert radians to degrees?  Degrees to radians?  Radians to fractions of a circle?  Degrees to fractions of a circle?}

\item{Can you draw a positive angle?  A negative angle?}

\item {Can you define and draw a right angle?}

\item{Can you define and find equivalent angles?  Do you understand why some angles are equivalent?}

\item{Can you define and draw complementary angles?  Supplementary angles?}

\item{Can you list and draw the different classifications of angles?}

\end{enumerate}
\subsection{Chapter 3}
\begin{enumerate}

\item{Can you define a trigonometric function?}

\item{Can you define Cartesean coordinate system?}

\item{Can you define a projection?  Do you understand why a trigonometric projection is at a right angle?}

\item{Can you define the sine of an angle?  The cosine? The tangent?}

\item{Can you define the reciprocal trigonometric functions:  the secant, the cosecant, the cotangent?}

\item{Can you define the trigonometric functions in terms of a right triangle instead of a projections?  Do you understand why you can?}

\end{enumerate}
\subsection{Chapter 4}
\begin{enumerate}


\item{Can you draw a complete unit circle with all the simple angles?}

\item{Can you list all the simple angles in degrees, radians, and fractions of a circle?}

\item{Can you list the sine of all the simple angles?  The cosine?  The tangent?}

\item{Do you understand why the tangent of $\frac{\pi}{2}$ is undefined?}

\end{enumerate}
\subsection{Chapter 5}
\begin{enumerate}


\item{Can you define a periodic function?  The period of a function?}

\item{Can you define phase?  Do you know the phase difference between sine and cosine?}

\item{Can you define asymptotes?}

\item{Can you draw and label a graph of all the trig functions?}\\


\end{enumerate}
\subsection{Chapter 6}
\begin{enumerate}

\item{Can you explain why trig. functions are periodic?  Why every angle has equivalent angles?  Can you find and draw equivalent angles for any angle?}

\item{Can you define even and odd functions?  Can you list which trig. functions are even and which are odd?}

\item{Can you define the Pythagorean Identity?  Explain its relationship to the Pythagorean Theorem?}

\item{Can you rearrange and simplify any expression composed of trig. functions?}

\end{enumerate}
\subsection{Chapter 7}
\begin{enumerate}

\item{Can you define an inverse function? Can you find the inverse of a function? Can you explain why not every function has an inverse function?} 

\item{Can you define a relation? Explain the difference between a function and a relation?}

\item{Can you define principle-value range?  Can you explain why inverse trig. functions have a principal value range?}

\item{Can you draw and label a graph of arc-sine, arc-cos, and arc-tangent?}

\end{enumerate}
\subsection{Chapter 8}
\begin{enumerate}

\item{Can you recite the formula for the sine of the sum of angles?  The cosine?  The tangent?}

\item{Can you derive the double-angle formulae for sine, cosine, and tangent?}

\item{Can you derive the half-angle formulae for sine, cosine, and tangent?}

\end{enumerate}
\subsection{Chapter 9}
\begin{enumerate}

\item{Can you draw and label rectangular coordinates and polar coordinates for a point?}

\item{Can you convert from polar coordinates to rectangular coordinates?}

\item{Can you convert from rectangular coordinates to polar coordinates?}

\end{enumerate}
\subsection{Chapter 10}
\begin{enumerate}

\item{Can you write the law of sines?  The law of cosines?}

\end{enumerate}
\subsection{Chapter 11}
\begin{enumerate}

\item{Can you define a real number?  An imaginary number?  A complex number?}

\item{Can you convert a complex number from rectangular form to polar form?  From polar form to rectangular form?}

\item{Can you add two complex numbers in rectangular form?  In polar form?}

\item{Can you subtract two complex numbers in rectangular form?  In polar form?}

\item{Can you multiply two complex numbers in rectangular form?  In polar form?}

\item{Can you divide two complex numbers in rectangular form?  In polar form?}

\end{enumerate}
\subsection{Chapter 12}
\begin{enumerate}

\item{What is Euler's formula?  Do you understand what it means with respect to rectangular and polar coordinates?}

\item{Can you use Euler's formula to derive different properties of sine and cosinem, such as $sin(-\theta) = -sin(\theta)$, or $cos(\theta) = \frac{1}{2}(e^{i\theta} + e^{-i\theta})$?}

\end{enumerate}
